Carbonate chemistry at Ny-Ålesund

### 2023-02-10 awipev-CO2.Rmd
Fixing bugs introduced by package updates

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
if (!require("latex2exp")) install.packages("latex2exp")
library("latex2exp")
if (!require("gtsummary")) install.packages("gtsummary")
library("gtsummary")
if (!require("todor")) install.packages("todor")
library("todor")
if (!require("ggpmisc")) install.packages("ggpmisc")
library("ggpmisc")
# todor::todor_file("awipev-CO2.Rmd")

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- d_all %>%
  dplyr::mutate(
    pco2_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_fb_ctd_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_ctd),
    fco2_fb = seacarb::p2fCO2(
      pCO2 = pco2_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_calc_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    fco2_calc_fb = seacarb::p2fCO2(
      pCO2 = pco2_calc_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_calc_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m)
) %>% 
  dplyr::select(sort(current_vars()))

@@

```{r Range per month, echo = FALSE, warning = FALSE}
range_month <-
  d_all %>%
  dplyr::select(montht, sal_fb_mix, temp_insitu_11m_mix, temp_insitu_11m, pHint_tot_sf_10m) %>% 
tbl_summary(by = montht,
            missing = "no",
            type = all_continuous() ~ "continuous2",
            statistic = all_continuous() ~ c(
              "{mean}",
              "{median}",
              "{p25}",
              "{p75}",
              "{min}",
              "{max}"),
            digits = list(sal_fb_mix ~ 1, temp_insitu_11m_mix ~ 1, temp_insitu_11m ~ 1, pHint_tot_sf_10m ~ 3))
range_month
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pCO2 NOT normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_remos = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_remos_ctd_temp = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_remos,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  )
# add at_remos and pco2_remos_ctd_temp to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, at_remos, pco2_remos_ctd_temp), by = "datetime") %>%
  dplyr::select(sort(current_vars()))
tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_ctd_temp)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group * tmp$montht))

@@

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_remos_ctd_temp_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-46,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at in situ temperature ($\\mu atm$)")) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 10,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 10,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface']<pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
pco2_delta_plt
# we look at pCO2 normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                #pHint_tot_sf,
                pco2_remos_ctd_temp,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(pco2_remos_4 = seacarb::pCO2insi(pCO2lab = pco2_remos_ctd_temp, Tlab = temp_insitu_ctd, SST = 4)
  )
# add pco2_remos_4 to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, pco2_remos_4), by = "datetime") %>%
  dplyr::select(sort(current_vars()))

@@

d_all <- dplyr::left_join(d_all, dplyr::select(atm_co2, datetime, co2_air), by = "datetime") %>% 
  dplyr::select(sort(current_vars()))

@@

# TODO: update this dataset
load(file = paste0(path, "meteo_awi/wind_nyalesund.RData")) 
meteo <- dplyr::tibble(windnyalesund) %>% 
  dplyr::slice_head(n = nrow(windnyalesund) - 1) %>% # remove last row which only contains NAs
  dplyr::mutate(datetime = as_datetime(paste(paste(Year, Month, Day, sep="-"), paste(Hour, "00", "00", sep="-"), sep = " ")),
         temp_air = `Temperature[degC]`,
         p_air = `Pressure[hPa]`,
         ws = `WindSpeed[m/s]`,
         wd = `WindDirection[deg]`
        ) %>%
  dplyr::select(datetime, temp_air, p_air, ws, wd)
d_all <- dplyr::left_join(d_all, meteo, by = "datetime") %>% 
  dplyr::select(sort(current_vars()))
```

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
p_pco2_fb_remos <- d_all %>% 
  dplyr::select(depth, monthd, pco2_fb_ctd_temp, pco2_remos_ctd_temp) %>% 
  dplyr::filter(depth < 4) %>% 
  drop_na() %>% 
  ggplot(aes(x = pco2_fb_ctd_temp, y = pco2_remos_ctd_temp, label = monthd)) +
  geom_point(col="blue", size = 1) +
  xlim(250, 400) + ylim(250, 400) +
  labs(title = "pCO2 at ctd temperature",
       x=latex2exp::TeX("$pCO_{2}$ 11 m $\\phantom{x}(\\mu atm)$"),
       y= latex2exp::TeX("$pCO_{2}$ REMOS at 0-4 m\\phantom{x}($\\mu atm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  ggpmisc::stat_ma_eq(mapping = use_label(c("R2", "eq"))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 11) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
#p_pco2_fb_remos
ggsave(filename = "figures/pco2_fb_remos.png", plot = p_pco2_fb_remos, width = 25, height = 25, units = "cm")

@@

```{r, fig.width=10, fig.height=10, echo=FALSE, message= FALSE, warning = FALSE}
# we calculate it from at_calc and pCO2
# nuts are not taken into consideration
tmp <- d_all %>% # just for this chunk
  dplyr::select(datetime, monthd, sal_fb_mix, temp_fb, pco2_fb, at_calc) %>% 
  tidyr::drop_na() %>% 
  dplyr::mutate(OmegaAragonite = seacarb::carb(
    flag = 24, var1 = pco2_fb, var2 = at_calc*1e-6, S=sal_fb_mix, T=temp_fb, 
    Patm=1, P=1.1, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T", b="u74")$OmegaAragonite)
#plot
size_point <- 0.3 # size of data points
alpha <- 1 #transparency
size_labs <- 6 # font size labels

ts_oa <- tmp %>%
  ggplot(aes(x = datetime, y = OmegaAragonite), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  ylim(0, 3) +
  labs(title="Aragonite saturation state (11 m)",x="", y=latex2exp::TeX("$\\Omega_{aragonite}$")) +
  Mytheme(size_labs = size_labs)
ts_oa
ggsave(filename = "figures/OmegaAragonite.png", plot = ts_oa, width = 15, height = 10, units = "cm")
```

### Whole time series plot

>>>>

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
if (!require("latex2exp")) install.packages("latex2exp")
library("latex2exp")
if (!require("gtsummary")) install.packages("gtsummary")
library("gtsummary")
if (!require("todor")) install.packages("todor")
library("todor")
if (!require("ggpmisc")) install.packages("ggpmisc")
library("ggpmisc")
if (!require("tidyselect")) install.packages("tidyselect")
library("tidyselect")
# todor::todor_file("awipev-CO2.Rmd")

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- d_all %>%
  dplyr::mutate(
    pco2_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_fb_ctd_temp = seacarb::pCO2insi(
      pCO2lab = pco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_ctd),
    fco2_fb = seacarb::p2fCO2(
      pCO2 = pco2_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    pco2_calc_fb_insi_temp = seacarb::pCO2insi(
      pCO2lab = pco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m),
    fco2_calc_fb = seacarb::p2fCO2(
      pCO2 = pco2_calc_fb,
      T = temp_fb,
      Patm = 1,
      P = 0),
    fco2_calc_fb_insi_temp = seacarb::fCO2insi(
      fCO2lab = fco2_calc_fb,
      Tlab = temp_fb,
      SST = temp_insitu_11m)
) %>% 
  dplyr::select(sort(everything()))

@@

```{r Range per month, echo = FALSE, warning = FALSE}
# TODO
# range_month <-
#   d_all %>%
#   dplyr::select(montht, sal_fb_mix, temp_insitu_11m_mix, temp_insitu_11m, pHint_tot_sf_10m) %>% 
# tbl_summary(by = montht,
#             missing = "no",
#             type = all_continuous() ~ "continuous2"),
#             statistic = all_continuous() ~ c(
#               "{mean}",
#               "{median}",
#               "{p25}",
#               "{p75}",
#               "{min}",
#               "{max}"
#               ),
#             digits = list(sal_fb_mix ~ 1, temp_insitu_11m_mix ~ 1, temp_insitu_11m ~ 1, pHint_tot_sf_10m ~ 3))
# range_month
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pCO2 NOT normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_remos = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_remos_ctd_temp = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_remos,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  )
# add at_remos and pco2_remos_ctd_temp to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, at_remos, pco2_remos_ctd_temp), by = "datetime") %>%
  dplyr::select(sort(everything()))
tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_ctd_temp)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group * tmp$montht))

@@

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_remos_ctd_temp_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-46,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at in situ temperature ($\\mu atm$)")) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 10,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']"
  ) +
  annotate(
    "text",
    x = 10,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface']<pCO['2 deep']"
  ) +
  Mytheme(size_labs = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
pco2_delta_plt
# we look at pCO2 normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                #pHint_tot_sf,
                pco2_remos_ctd_temp,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(pco2_remos_4 = seacarb::pCO2insi(pCO2lab = pco2_remos_ctd_temp, Tlab = temp_insitu_ctd, SST = 4)
  )
# add pco2_remos_4 to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, pco2_remos_4), by = "datetime") %>%
  dplyr::select(sort(everything()))

@@

d_all <- dplyr::left_join(d_all, dplyr::select(atm_co2, datetime, co2_air), by = "datetime") %>% 
  dplyr::select(sort(everything()))

@@

# TODO: update this dataset
load(file = paste0(path, "meteo_awi/wind_nyalesund.RData")) 
meteo <- dplyr::tibble(windnyalesund) %>% 
  dplyr::slice_head(n = nrow(windnyalesund) - 1) %>% # remove last row which only contains NAs
  dplyr::mutate(datetime = as_datetime(paste(paste(Year, Month, Day, sep="-"), paste(Hour, "00", "00", sep="-"), sep = " ")),
         temp_air = `Temperature[degC]`,
         p_air = `Pressure[hPa]`,
         ws = `WindSpeed[m/s]`,
         wd = `WindDirection[deg]`
        ) %>%
  dplyr::select(datetime, temp_air, p_air, ws, wd)
d_all <- dplyr::left_join(d_all, meteo, by = "datetime") %>% 
  dplyr::select(sort(everything()))
```

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
p_pco2_fb_remos <- d_all %>% 
  dplyr::select(depth, monthd, pco2_fb_ctd_temp, pco2_remos_ctd_temp) %>% 
  dplyr::filter(depth < 4) %>% 
  dplyr::filter(if_all(everything(), is.finite)) %>% 
  ggplot(aes(x = pco2_fb_ctd_temp, y = pco2_remos_ctd_temp, label = monthd)) +
  geom_point(col="blue", size = 1) +
  xlim(250, 400) + ylim(250, 400) +
  labs(title = "pCO2 at ctd temperature",
       x=latex2exp::TeX("$pCO_{2}$ 11 m $\\phantom{x}(\\mu atm)$"),
       y= latex2exp::TeX("$pCO_{2}$ REMOS at 0-4 m\\phantom{x}($\\mu atm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  ggpmisc::stat_ma_eq(mapping = use_label(c("R2", "eq"))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 11) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
#p_pco2_fb_remos
ggsave(filename = "figures/pco2_fb_remos.png", plot = p_pco2_fb_remos, width = 25, height = 25, units = "cm")

@@

```{r, fig.width=10, fig.height=10, echo=FALSE, message= FALSE, warning = FALSE}
# we calculate it from at_calc and pCO2
# nuts are not taken into consideration
tmp <- d_all %>% # just for this chunk
  dplyr::select(datetime, monthd, sal_fb_mix, temp_fb, pco2_fb, at_calc) %>% 
  tidyr::drop_na() %>% 
  dplyr::mutate(OmegaAragonite = seacarb::carb(
    flag = 24, var1 = pco2_fb, var2 = at_calc*1e-6, S=sal_fb_mix, T=temp_fb, 
    Patm=1, P=1.1, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T", b="u74")$OmegaAragonite)
summary(tmp$OmegaAragonite)
#plot
size_point <- 0.3 # size of data points
alpha <- 1 #transparency
size_labs <- 6 # font size labels

ts_oa <- tmp %>%
  ggplot(aes(x = datetime, y = OmegaAragonite), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  ylim(0, 3) +
  labs(title="Aragonite saturation state (11 m)",x="", y=latex2exp::TeX("$\\Omega_{aragonite}$")) +
  Mytheme(size_labs = size_labs)
ggsave(filename = "figures/OmegaAragonite.png", plot = ts_oa, width = 15, height = 10, units = "cm")
```
### Whole time series plot
